---
title: "grade analysis"
format: html
---
```{r}
#| message: false

here::i_am("r-101-grade-analysis.Rproj")

library(here)
library(ggplot2)
library(tidyr)
library(dplyr)
library(vroom)
```

## Question 1

```{r}
#| message: false
grades <- vroom(here("grades.csv"))
```

## Question 2
```{r}
library(knitr)
```

```{r}
grades |> summarise("min"=min(Exam, na.rm=TRUE), "max"= max(Exam, na.rm=TRUE), "median"= median(Exam, na.rm=TRUE), "mean"= mean(Exam, na.rm=TRUE)) |> knitr::kable()
```
## Question 3

```{r}
num_absent <- grades |> filter(is.na(Exam))
```

There were  `r nrow(num_absent)` absent students. 

## Question 4

### Alternative 1:

```{r}
grades_new <- grades |> filter(!is.na(Exam))
```
```{r}
grade_ <- grades |> filter(!is.na(Exam)) |> group_by(Exam) |> summarise(N=n())
```

```{r}
ggplot(grade_, aes(x=Exam, y=N)) + geom_col()
```
### Alternative 2
```{r}
ggplot(grades_new, aes(x=Exam)) + geom_boxplot()
```
## Question 5

```{r}
grades |> group_by(Group) |> summarise(N=n()) |> knitr::kable()
```
## Question 6

```{r}
count_group <- grades |> group_by(Group) |> summarise(N=n())
```

```{r}
ggplot(count_group, aes(x=Group, y=N)) + geom_col()
```

## Question 7

Represent graphically the distribution of the grades at the Exam conditioned by the group. Test at least two different representations.

### Alternative 1
```{r}
ggplot(grades_new, aes(x=Exam, fill=Group)) + geom_bar(position="fill")
```
### Alternative 2

```{r}
ggplot(grades_new, aes(y=Exam, fill= Group)) + geom_boxplot()
```
## Question 8

```{r}
absent_pergroup <- grades |> mutate(absent= case_when(!is.na(Exam)~F,is.na(Exam)~T)) |>group_by(Group) |> summarise(sum= sum(absent)) 

absent_pergroup|> knitr::kable(caption="number of absent students per group")
```

## Question 9

```{r}
ggplot(absent_pergroup, aes(x=Group, y=sum)) + geom_col()
```

## Question 10

```{r}
columns <- grades |> select(is.numeric, -Id) |> names()
grades[columns] <-  round(grades[columns],2)
```

```{r}
grades_longer <- grades |> pivot_longer(-c("Id", "Group"),names_to= "name", values_to= "value") 
```

## Question 11

```{r}
absence <- grades_longer |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T)) |> group_by(Id) |> summarise(sum= sum(absent)) 

absence |> knitr::kable(caption = "number of missing grades per student") |> head(n = 10)
```

## Question 12

```{r}
ggplot(absence, aes(x= sum)) + geom_bar()
```

## Question 13

```{r}
grades_longer |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T))|> filter(name=="Exam") |> group_by(Group) |> summarise(sum= sum(absent)) 
```

```{r}
library(stringr)
```

## Question 14

```{r}
grades_longer_new <- grades_longer|> filter(str_starts(name, "Online")) |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T)) |> group_by(Id) |> summarise(sum= sum(absent)) 
grades_longer_new |> knitr::kable()
```

## Question 15

### Alternative 1: row-wise
```{r}
grades_NEW <- grades |> rowwise() |> mutate(absent= list(case_when(!is.na(c_across(starts_with("Online")))~F,is.na(c_across(starts_with("Online")))~T))) |> mutate(absent= sum(absent))
 
```
```{r}
ggplot(grades_NEW, aes(x=Exam, fill=factor(absent))) + geom_bar(position="fill")
```

### Alternative 2

```{r}
grades_new <- inner_join(grades, grades_longer_new, by=join_by("Id"))
```
```{r}
ggplot(grades_new, aes(x=Exam, fill=factor(sum))) + geom_bar(position="fill")
```

## Question 16

Create a table with two columns: Id to identify students and Missed with value TRUE when the student miss at least one MCQ_xx grade and FALSE when they miss no grade.

```{r}
table <- grades |> rowwise() |> mutate(absent_mcq= list(case_when(!is.na(c_across(starts_with("MCQ")))~F,is.na(c_across(starts_with("MCQ")))~T))) |> mutate(absent_mcq= sum(absent_mcq)) |> mutate(Missed=case_when(absent_mcq > 0 ~ T, absent_mcq==0 ~ F)) |> select(Id, Missed)
```

## Question 17

```{r}
grades <- grades |> rowwise() |> mutate(absent_mcq= list(case_when(!is.na(c_across(starts_with("MCQ")))~F,is.na(c_across(starts_with("MCQ")))~T))) |> mutate(absent_mcq= sum(absent_mcq)) |> mutate(Missed=case_when(absent_mcq > 0 ~ T, absent_mcq==0 ~ F)) 

Pmissed <- grades |> group_by(Group) |> summarise(P_missed=sum(Missed)/n())
```

## Question 18

Represent graphically the average grade at the Exam per group as a function of the
percentage of missed MCQ_xx grade as defined in the previous question. The simplest way to obtain the
data is to use inner_join to merge two intermediate tables (the one obtained in the previous question
and one giving the average grades per group).

```{r}
exam_grades <- grades |> group_by(Group) |> summarise(mean= mean(Exam,  na.rm=TRUE))
table2 <- inner_join(Pmissed, exam_grades)
```
```{r}
ggplot(table2, aes(x=P_missed, y=mean)) + geom_line()
```
