---
title: "grade analysis"
format: html
---

```{r}
#| message: false

here::i_am("r-101-grade-analysis.Rproj")

library(here)
library(ggplot2)
library(tidyr)
library(dplyr)
library(vroom)
```

## Question 1

```{r}
#| message: false
grades <- vroom(here("grades.csv"))
```

## Question 2

```{r}
library(knitr)
```

```{r}
grades |> summarise("min"=min(Exam, na.rm=TRUE), "max"= max(Exam, na.rm=TRUE), "median"= median(Exam, na.rm=TRUE), "mean"= mean(Exam, na.rm=TRUE)) |> knitr::kable(caption= "statistics of Exam")
```

## Question 3

```{r}
num_absent <- grades |> filter(is.na(Exam))
```

There were `r nrow(num_absent)` absent students.

## Question 4

### Alternative 1:

```{r}
grade_ <- grades |> filter(!is.na(Exam)) |> group_by(Exam) |> summarise(N=n())
```

```{r distribution of exam (1) }
ggplot(grade_, aes(x=Exam, y=N)) + geom_col() + ylab("count") + ggtitle("distribution of Exam")
```

### Alternative 2

```{r}
grades_new <- grades |> filter(!is.na(Exam))
```

```{r distribution of exam (2)}
ggplot(grades_new, aes(y=Exam)) + geom_boxplot() + ggtitle("distribution of Exam")
```

## Question 5

```{r}
grades |> group_by(Group) |> summarise(N=n()) |> knitr::kable(caption = "repartition of students in each group")
```

## Question 6

```{r}
count_group <- grades |> group_by(Group) |> summarise(N=n())
```

```{r distribution of people in each group}
ggplot(count_group, aes(x=Group, y=N)) + geom_col() + ylab("count") + ggtitle("distribution of Exam") + coord_flip()
```

Here we used coord_flip to make the Group-axis more readable.

## Question 7

### Alternative 1:

```{r distribution of Exam conditioned on Group (1)}
ggplot(grades_new, aes(x=Exam)) + geom_density() + facet_wrap(~Group) + ggtitle("distribution of Exam conditioned on Group")
```

### Alternative 2:

```{r distribution of Exam conditioned on Group (2)}
ggplot(grades_new, aes(y=Exam,x= Group)) + geom_boxplot() + xlab("Group") + ggtitle("distribution of Exam conditioned by group") + ggtitle("distribution of Exam conditioned on Group")+ coord_flip()
```

## Question 8

```{r}
absent_pergroup <- grades |> mutate(absent= case_when(!is.na(Exam)~F,is.na(Exam)~T)) |> group_by(Group) |> summarise(absence= sum(absent)) 

absent_pergroup|> knitr::kable(caption="number of absent students per group")
```

## Question 9

```{r repartion of absence per group}
ggplot(absent_pergroup, aes(x=Group, y=absence)) + geom_col() + ylab("absence") + ggtitle("repartion of absence per group") + coord_flip()
```

## Question 10

```{r}
grades_longer <- grades |> pivot_longer(-c("Id", "Group"),names_to= "name", values_to= "value") 
grades_longer["value"] <- round(grades_longer["value"],2)
```

## Question 11

```{r}
absence <- grades_longer |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T)) |> group_by(Id) |> summarise(sum= sum(absent)) 

absence|> head(n = 10) |> knitr::kable(caption = "number of missing grades per student") 
```

## Question 12

```{r distribution of the number of missing grades per student}
ggplot(absence, aes(x= sum)) + geom_bar() + xlab("number of absence") + ylab("count") + ggtitle("distribution of the number of missing grades per student")
```

## Question 13

```{r}
grades_longer |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T))|> filter(name=="Exam") |> group_by(Group) |> summarise(absences= sum(absent))  |> knitr::kable()
```

```{r}
library(stringr)
```

## Question 14

```{r}
absence_online <- grades_longer|> filter(str_starts(name, "Online")) |> mutate(absent= case_when(!is.na(value)~F,is.na(value)~T)) |> group_by(Id) |> summarise(absence= sum(absent)) 

absence_online |> head(n=10) |> knitr::kable()
```

## Question 15

### Alternative 1: row-wise

```{r}
grades_NEW <- grades |> rowwise() |> mutate(absent= list(case_when(!is.na(c_across(starts_with("Online")))~F,is.na(c_across(starts_with("Online")))~T))) |> mutate(absent= sum(absent)) |> filter(!is.na(Exam)) 
```

```{r distribution of the grades at the Exam conditioned by missing grades for online tests (1)}
ggplot(grades_NEW, aes(x=Exam)) + geom_density() + facet_wrap(~absent) + ggtitle("distribution of the grades at the Exam conditioned by missing grades for online tests")
```
### Alternative 2:

```{r distribution of the grades at the Exam conditioned by missing grades for online tests (2)}
inner_join(grades, absence_online, by=join_by("Id")) |> filter(!is.na(Exam)) |> ggplot(aes(x=Exam)) + geom_density() + facet_wrap(~absence) + ggtitle("distribution of the grades at the Exam conditioned by missing grades for online tests")
```

## Question 16

```{r}
grades <- grades |> rowwise() |> mutate(absent_mcq= list(case_when(!is.na(c_across(starts_with("MCQ")))~F,is.na(c_across(starts_with("MCQ")))~T))) |> mutate(absent_mcq= sum(absent_mcq)) |> mutate(Missed=case_when(absent_mcq > 0 ~ T, absent_mcq==0 ~ F)) 
```

```{r}
grades |> select(Id, Missed) |> head(n=10) |> knitr::kable()
```

## Question 17

```{r}
Pmissed <- grades |> group_by(Group) |> summarise(P_missed=sum(Missed)/n())

Pmissed |> head(n=10) |> knitr::kable()
```

## Question 18

```{r}
#| message: false

exam_grades <- grades |> group_by(Group) |> summarise(mean= mean(Exam,  na.rm=TRUE))
table <- inner_join(Pmissed, exam_grades)
```

```{r average  Exam grade per group as a function of percentage of missed MCQ}
ggplot(table, aes(x=P_missed, y=mean)) + geom_point() + ggtitle("average  Exam grade per group as a function of percentage of missed MCQ")
```
