---
title: "grade analysis"
format: html
---
```{r}
#| message: false

here::i_am("r-101-grade-analysis.Rproj")

library(here)
library(ggplot2)
library(tidyr)
library(dplyr)
library(vroom)
```

## Question 1

```{r}
#| message: false
grades <- vroom(here("grades.csv"))
```

## Question 2
```{r}
library(knitr)
```


```{r}
knitr::kable(grades |> summary("min"=min(Exam), "max"= max(Exam), "median"= median(Exam), "mean"= mean(Exam)))
```
```{r}
grades |> summarise("min"=min(Exam, na.rm=TRUE), "max"= max(Exam, na.rm=TRUE), "median"= median(Exam, na.rm=TRUE), "mean"= mean(Exam, na.rm=TRUE)) |> knitr::kable()
```
## Question 3

```{r}
num_absent <- grades |> filter(is.na(Exam))
```

There were  `r nrow(num_absent)` absent students. 

## Question 4
Represent graphically in an adapted way the distribution of grades at the Exam. Make sure to adapt the code to avoid any error or warning from ggplot2.

```{r}
grades_new <- grades |> filter(!is.na(Exam))
```
```{r}
grade_ <- grades |> filter(!is.na(Exam)) |> group_by(Exam) |> summarise(N=n())
```

```{r}
ggplot(grade_, aes(x=Exam, y=N)) + geom_col()
```
```{r}
ggplot(grades_new, aes(x=Exam)) + geom_boxplot()
```
## Question 5

Compute the number of students in each Group and include the results in your document as a table

```{r}
grades |> group_by(Group) |> summarise(N=n()) |> knitr::kable()
```
## Question 6

Represent the same information as in the previous question in a graphical way
```{r}
count_group <- grades |> group_by(Group) |> summarise(N=n())
```

```{r}
ggplot(count_group, aes(x=Group, y=N)) + geom_col()
```

## Question 7

Represent graphically the distribution of the grades at the Exam conditioned by the group. Test at least two different representations.

### Alternative 1
```{r}
ggplot(grades, aes(x=Exam, fill=Group)) + geom_bar(position="fill")
```
### Alternative 2

```{r}
ggplot(grades_new, aes(y=Exam, fill= Group)) + geom_boxplot()
```
## Question 8

Compute the number of students who missed the Exam in each Group (i.e. students whose Exam is NA). Beware that this number can be zero in some groups and this can induce difficulties with
group_by. A way to circumvent this problem is to note that the sum of a vector of TRUE/FALSE values
is exactly the number of times TRUE values that appear in the vector. For instance

```{r}
grades |> mutate(absent= case_when(!is.na(Exam)~F,is.na(Exam)~T))
```
```{r}
grades |> mutate(absent= case_when(!is.na(Exam)~F,is.na(Exam)~T)) |>group_by(Group) |> summarise(sum= sum(absent))
```

